\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-06-04 paper/main.tex
.

\abstract{\textbf{Motivation:} \textit{In silico} drug--target
  interaction (DTI) prediction is important for drug discovery and
  drug repurposing.  Approaches to predict DTIs can proceed
  indirectly, top-down, using phenotypic effects of drugs to identify
  potential drug targets, or they can be direct, bottom-up and use
  molecular information to directly predict binding potentials.  Both
  approaches can be combined with information about interaction
  networks. \\
  \textbf{Results:} We developed \name{} as a computational method
  that combines molecular features and ontology-encoded phenotypic
  effects of drugs with protein--protein interaction networks, and
  uses a graph convolutional neural network to predict DTIs.  We
  demonstrate that drug effect features can exploit information in the
  interaction network whereas molecular features do not.  \name{} is
  designed to predict candidate drugs for a given protein; we use this
  formulation to show that common DTI datasets contain intrinsic
  biases with major effects on performance evaluation and comparison
  of DTI prediction methods. Using a modified evaluation scheme, we
  demonstrate that \name{} improves significantly over state of the
  art DTI prediction methods. \\
  \textbf{Availability:} \name{} source code and data necessary to
  reproduce results are freely available at
  \url{https://github.com/THinnerichs/DTI-VOODOO}. \\
  \textbf{Contact:}
  \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}
  \\
  \textbf{Supplementary information:}Supplementary data are available
  at \textit{Bioinformatics} online.}

>>>>

\abstract{\textbf{Motivation:} \textit{In silico} drug--target
  interaction (DTI) prediction is important for drug discovery and
  drug repurposing.  Approaches to predict DTIs can proceed
  indirectly, top-down, using phenotypic effects of drugs to identify
  potential drug targets, or they can be direct, bottom-up and use
  molecular information to directly predict binding affinities.  Both
  approaches can be combined with information about interaction
  networks. \\
  \textbf{Results:} We developed \name{} as a computational method
  that combines molecular features and ontology-encoded phenotypic
  effects of drugs with protein--protein interaction networks, and
  uses a graph convolutional neural network to predict DTIs.  We
  demonstrate that drug effect features can exploit information in the
  interaction network whereas molecular features do not.  \name{} is
  designed to predict candidate drugs for a given protein; we use this
  formulation to show that common DTI datasets contain intrinsic
  biases with major effects on performance evaluation and comparison
  of DTI prediction methods. Using a modified evaluation scheme, we
  demonstrate that \name{} improves significantly over state of the
  art DTI prediction methods. \\
  \textbf{Availability:} \name{} source code and data necessary to
  reproduce results are freely available at
  \url{https://github.com/THinnerichs/DTI-VOODOO}. \\
  \textbf{Contact:}
  \href{tilman.hinnerichs@kaust.edu.sa}{tilman.hinnerichs@kaust.edu.sa}
  \\
  \textbf{Supplementary information:}Supplementary data are available
  at \textit{Bioinformatics} online.}
